Course Project 1 Reproducible Research
=========================================
Jersai


## Loading and preprocessing the data 


* Load the data
```{r loadingdata}
library(ggplot2)
activity <- read.csv("activity.csv")
```


##  What is mean total number of steps taken per day?


* Calculate the total number of steps taken per day
```{r numberofsteps}
stepsPerDay <- aggregate(steps ~ date, activity, sum, na.rm=TRUE)
stepsPerDay
```


* Make a histogram of the total number of steps taken each day
```{r histogram}
hist(stepsPerDay$steps, breaks= seq(0,25000,by=2500), ylim=c(0,20), col=5, main="Total number of steps taken each day")
```


* Calculate and report the mean and median of the total number of steps taken per day
```{r meanmedian}
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```


##  What is the average daily activity pattern?


*  Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across al days (y-axis)
```{r timeseries}
stepsPerInterval <- aggregate(steps ~ interval, activity, mean, na.rm=TRUE)
g <- ggplot(data= stepsPerInterval, aes(x=interval, y=steps))
g <- g + geom_line() + labs(x="5-minute interval", y="steps",title="Average daily activity pattern")
print(g)
```


* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxsteps}
stepsPerInterval[which.max(stepsPerInterval$steps), ]$interval
```


## Imputing missing values


* Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r missingvalues}
NASteps <- sum(is.na(activity$steps))
NASteps
```


* Devise a strategy for filling in all of the missing values in the dataset. In this case, we will be evaluating, using skewness function, whether the variable steps has a symmetric distribution or not. If skewness is different than 0, missing values will be imputed using the median, otherwise the mean will be used.
```{r strategy}
library(moments)
skewness(activity$steps,na.rm=TRUE) 
medianSteps <- median(activity$steps,na.rm=TRUE)
```


* Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r newdataset}
activity2 <- activity
activity2$steps[is.na(activity2$steps)] <- medianSteps
```


* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
``` {r histogram2}
stepsPerDay2 <- aggregate(steps ~ date, activity2, sum, na.rm=TRUE)
hist(stepsPerDay2$steps, breaks= seq(0,25000,by=2500), ylim=c(0,20), col=5, main="Total number of steps taken each day")
```


## Are there differences in activity patterns between weekdays and weekends?

* Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
``` {r factorvar}
activity2$date <- as.Date(activity2$date)
activity2$day <- weekdays(activity2$date)
for (i in 1:nrow(activity2)) {
    if (activity2[i,]$day %in% c("Saturday","Sunday")) {
        activity2[i,]$day<-"weekend"
    }
    else{
        activity2[i,]$day<-"weekday"
    }
}
```


* Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r panelplot}
stepsPerInterval2 <- aggregate(steps ~ interval + day, activity2, mean, na.rm=TRUE)
g <- ggplot(data= stepsPerInterval2, aes(x=interval, y=steps))
g <- g + geom_line() + facet_grid(day~.) + labs(x="5-minute interval", y="steps",title="Average daily activity pattern across all weekdays or weekend days")
print(g)
```